---
title: "Bacteria Assessment"
author: "Andrew Feltmann"
date: "8/23/2022"
output: html_document
---


```{r setup, include=FALSE}
library(RODBC)
library(stringr)
library(foreign)
library(dplyr)
library(tidyr)
library(zoo)
library(lubridate)
library(psych)
```

```{r include = F}
options(scipen = 999)
```

# Data read

Set up the database connection and ead in the Bacteria data

```{r}
con <- RODBC::odbcConnectAccess2007("E:/2022 Integrated Report/Data/Phase 3 QC/DB Back-Up/PHASE 3 DATASET BACKUP 2022-11-16 - Copy.accdb")



primary_data <- sqlFetch(con, "Bact Data")

RODBC::odbcClose(con)
```

# Processing

Clean up variable names and data representations.

```{r}
colnames(primary_data)[3] <- "CollectionDate"
colnames(primary_data)[1] <- "AU"
colnames(primary_data)[7] <- "ORW"

primary_data$CollectionDate <- mdy(primary_data$CollectionDate)
primary_data$ParameterName <- str_trim(primary_data$ParameterName)
```

Read in the distribution QC failures and filter the bacteria data to remove them from what we will be processing.

```{r}
QC_fail_aus <- read.csv("E:/2022 Integrated Report/Data/Phase 3 QC/WQAR Setup/Final_Dist_Decisions_sample_year fails added.csv")

QC_fail_aus <- QC_fail_aus %>%
        separate(Pollutant.Pairs, into = c("AU", "Param"), sep = "-")%>%
        filter(Param == "E. coli")

primary_data <- primary_data %>%
        filter(!AU %in% QC_fail_aus$AU)
```

Calculate the Attainments and Failures, each source is for a methodology per the assessment methodology. Refer to the AM for an explanation of how Individual Samples (IS) and Geometric mean (GM) are calculated

```{r}
source("E:/2022 Integrated Report/Data Assessments/BACTERIA ASSESSMENTS/scripts/Bact_gmassess_function.R")
source("E:/2022 Integrated Report/Data Assessments/BACTERIA ASSESSMENTS/scripts/Bact_isassess_function.R")
```

```{r}
bact_assessment_final <- rbind(bact_assessment_is,
                               bact_assessment_gm)

bact_assessment_final <- bact_assessment_final %>%
        separate(bact_pairing, c("AU", "Season", "SiteType", "ORW"), sep = "-")
```

With the completion of the code, two csv's are generated:

* **"BACTERIA ASSESSMENTS"** is all of the assessments for both methodologies with their respective ATTAIN/FAIL designations
* **"BACTERIA IS ASSESSMENT SUMMARY"** is a QC document for checking the exceedance rates for AUs that were assessed under the individual sample methodology. 

```{r}
setwd("E:/2022 Integrated Report/Data Assessments/BACTERIA ASSESSMENTS/output")

write.csv(bact_assessment_final, 
          "BACTERIA ASSESSMENTS.csv", 
          row.names = F)

write.csv(bact_assessment_is_sum, 
          "BACTERIA IS ASSESSMENT SUMMARY.csv",
          row.names = F)
```
